# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r echo=TRUE}
library(reshape2)

unzip("activity.zip")
activity=read.csv("activity.csv")
summary(activity)
by_days_steps_list=lapply(split(activity,activity$date),  
                           function (x) sum(x[,c(1)])
                          )
#convert list of dataframes into a dataframe
by_days_steps_df = melt(by_days_steps_list)
names(by_days_steps_df)=c("Steps","Date")
```



## What is mean total number of steps taken per day?
```{r}

by_days_steps_df$Date = as.POSIXct(by_days_steps_df$Date,  format="%Y-%m-%d")
r=as.POSIXct(round(range(by_days_steps_df$Date),"hours"))
plot(by_days_steps_df$Date, by_days_steps_df$Steps, type="h", xaxt="n")
axis.POSIXct(side=1, at=seq(r[1], r[2], by="day"), format="%Y-%m-%d")
summary(by_days_steps_df$Steps)
```

The mean total number of steps taken per day is `r mean(by_days_steps_df$Steps, na.rm=TRUE)`.   
Median is `r median(by_days_steps_df$Steps, na.rm=TRUE)`

## What is the average daily activity pattern?
```{r}
 activity$date = as.POSIXct(paste(activity$date, sprintf("%04d", activity$interval)),  format="%Y-%m-%d %H%M")
r=as.POSIXct(round(range(activity$date),"hours"))
plot(activity$date, activity$steps, type="l", xaxt="n")
axis.POSIXct(side=1, at=seq(r[1], r[2], by="day"), format="%Y-%m-%d")

```





## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
