# Reproducible Research: Peer Assessment 1



## Loading and preprocessing the data
```{r echo=TRUE}
library(reshape2)
library(timeDate)
library(lattice)

unzip("activity.zip")
activity=read.csv("activity.csv")
summary(activity)
by_days_steps_list=lapply(split(activity,activity$date),  
                           function (x) sum(x[,c(1)])
                          )
#convert list of dataframes into a dataframe
by_days_steps_df = melt(by_days_steps_list)
names(by_days_steps_df)=c("Steps","Date")
```



## What is mean total number of steps taken per day?
```{r echo=TRUE}
hist( by_days_steps_df$Steps, breaks=40)

```

```{r fig.width=17, fig.height=6 echo=TRUE}
activity$date = as.POSIXct(paste(activity$date, sprintf("%04d", activity$interval)),  format="%Y-%m-%d %H%M")
r=as.POSIXct(round(range(activity$date),"hours"))
plot(as.POSIXct(by_days_steps_df$Date), by_days_steps_df$Steps,type="h",xaxt="n", xlab="", ylab="Steps")
axis.POSIXct(side=1, at=seq(r[1], r[2], by="day"), format="%Y-%m-%d", las=2)
summary(by_days_steps_df$Steps)
```

The mean total number of steps taken per day is `r mean(by_days_steps_df$Steps, na.rm=TRUE)`.   
Median is `r median(by_days_steps_df$Steps, na.rm=TRUE)`

## What is the average daily activity pattern?
```{r echo=TRUE}

by_interval_avg_steps_list=lapply(split(activity,activity$interval ),  
                           function (x) mean(x[,c(1)], na.rm=TRUE) )
                          
#convert list of dataframes into a dataframe
by_interval_avg_steps_list_df = melt(by_interval_avg_steps_list)
names(by_interval_avg_steps_list_df)=c("Steps","Interval")

plot(by_interval_avg_steps_list_df$Interval, by_interval_avg_steps_list_df$Steps, type="l")
```


The 5-minute pattern that contains the max number of steps on average across all the days in the data set is `r by_interval_avg_steps_list_df[by_interval_avg_steps_list_df$Steps==max(by_interval_avg_steps_list_df$Steps), 2] `


## Imputing missing values
Number of missing entries (for step column) is `r sum(is.na(activity$steps))`  

```{r echo=TRUE}
#use the mean of 5 minute interval to replace missing data in steps columns
activity=read.csv("activity.csv")
by_interval_split_list = lapply(split(activity,activity$interval ),
        function (x)   { 
          x$steps[is.na(x$steps)] <- mean(x[,c(1)], na.rm=TRUE) 
          x
          } 
          )   
full_activity= unsplit(by_interval_split_list, f=activity$interval) 
summary(full_activity)

```

```{r echo=TRUE}
# a histogram of the total number of steps taken each day
by_days_steps_list=lapply(split(full_activity,full_activity$date),  
                           function (x) sum(x[,c(1)])
                          )
#convert list of dataframes into a dataframe
by_days_steps_df = melt(by_days_steps_list)
names(by_days_steps_df)=c("Steps","Date")

hist( by_days_steps_df$Steps, breaks=40)
```
```{r echo=TRUE fig.width=17, fig.height=6}
full_activity$date = as.POSIXct(paste(full_activity$date, sprintf("%04d", full_activity$interval)),  format="%Y-%m-%d %H%M")
r=as.POSIXct(round(range(full_activity$date),"hours"))
plot(as.POSIXct(by_days_steps_df$Date), by_days_steps_df$Steps,type="h",xaxt="n", xlab="", ylab="Steps")
axis.POSIXct(side=1, at=seq(r[1], r[2], by="day"), format="%Y-%m-%d", las=2)
summary(by_days_steps_df$Steps)
```
The mean total number of steps taken per day is `r mean(by_days_steps_df$Steps, na.rm=TRUE)`.   
Median is `r median(by_days_steps_df$Steps, na.rm=TRUE)`
No diffrenece comparing to a raw dataset with NAs.

## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}
#find weekdays/weekends
#full_activity$day=isWeekday(full_activity$date)
#full_activity$day[full_activity$day==TRUE] = "Weekday"
#full_activity$day[full_activity$day==FALSE] = "Weekend"
#full_activity$day=factor(full_activity$day)

#better code
full_activity$day=factor(ifelse(isWeekday(full_activity$date),"Weekday", "Weekend"))
by_daytype_full_activity_list=split(full_activity, full_activity$day)
# by_interval_avg_steps_list=c(lapply(split(by_daytype_full_activity_list[["Weekday"]],by_daytype_full_activity_list[["Weekday"]]$interval ),  
#                            function (x) list(Steps=mean(x[,c(1)], na.rm=TRUE), Day="Weekday")
#                            ),
#                            lapply(split(by_daytype_full_activity_list[["Weekend"]],by_daytype_full_activity_list[["Weekend"]]$interval ),  
#                            function (x) list(mean(x[,c(1)], na.rm=TRUE), "Weekend")
#                            ) )

by_interval_avg_steps_weekday_list=lapply(split(by_daytype_full_activity_list[["Weekday"]],by_daytype_full_activity_list[["Weekday"]]$interval ),  
                           function (x) list(Steps=mean(x[,c(1)], na.rm=TRUE), Day="Weekday")
                           )
                          
by_interval_avg_steps_weekend_list=lapply(split(by_daytype_full_activity_list[["Weekend"]],by_daytype_full_activity_list[["Weekend"]]$interval ),  
                           function (x) list(Steps=mean(x[,c(1)], na.rm=TRUE), Day="Weekend")
                           )

#convert list into a dataframe
by_interval_avg_steps_df = do.call(rbind.data.frame,by_interval_avg_steps_weekday_list)
by_interval_avg_steps_df$Interval=rownames(by_interval_avg_steps_df)
by_interval_avg_steps_df$Interval=as.numeric(by_interval_avg_steps_df$Interval)

by_interval_avg_steps_weekend_df = do.call(rbind.data.frame,by_interval_avg_steps_weekend_list)
by_interval_avg_steps_weekend_df$Interval=rownames(by_interval_avg_steps_weekend_df)
by_interval_avg_steps_weekend_df$Interval=as.numeric(by_interval_avg_steps_weekend_df$Interval)

by_interval_avg_steps_df=rbind(by_interval_avg_steps_df, by_interval_avg_steps_weekend_df)

#plot(by_interval_avg_steps_df$Steps ,by_interval_avg_steps_df$Interval, type="l")
xyplot(by_interval_avg_steps_df$Steps ~ by_interval_avg_steps_df$Interval|by_interval_avg_steps_df$Day, data=by_interval_avg_steps_df, layout=c(1,2), type="l")

#xyplot(full_activity$steps ~ full_activity$date|full_activity$day, data=full_activity, layout=c(1,2), type="l")



```

